---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# SplineHazardRegression

<!-- badges: start -->
<!-- badges: end -->

There is no R software available for the direct estimation of hazards with uncertainty estimation (confidence intervals).

The primary goal of SplineHazardRegression is to make available the methods 
for flexible estimation of hazards using (cubic) b-splines published in 
    Philip S. Rosenberg. "Hazard Function Estimation Using B-Splines"
    In Biometrics, Vol. 51, No. 3 (Sep., 1995), pp. 874-887
    https://doi.org/10.2307/2532989
The input data is time-to-event data (e.g. time to death), possibly right-censored and with late entries (both meaning that patients are followed-up for unequal times).

The package also allows the flexible estimation of the cumulative hazard and cumulative survival functions, as well as the compuatation of aggregate measures for those (average, median, interquartile range, etc). 

Different methods for the automatic selection of knots and for variance estimation are implemented.

## Installation

You can install the development version of SplineHazardRegression from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("fespuny/SplineHazardRegression")
```

## Example

This is a basic example which shows you how to simulate time-to-event data and fit a hazard function, deriving then cumulative hazard and survival estimates.

### Data Simulation

```{r simulation}
library(SplineHazardRegression)
## simulation parameters
   knots = c(0, 1, 3, 6, 10, NA, NA)
   betac = 1 * c(0.05, 0.05, 0.05, 0.05, 0.40, 0.1, 0.05)
   HParm = data.frame(knots, betac) # 'A Simple B-Spline'
   cll = c(0, 5)
   cup = c(5, 10)
   cih = c(0.0125, 0.025)
   CParm = data.frame(cll, cup, cih) # 'Light Censoring'
## calculate simulation true hazard and censoring distributions
   INPUTS = etsim_inputs( HParam=HParm, CParam=CParm, SampleSize = 201 )
## simulate time-to-event data using true distribution
   SimDat = etsim(INPUTS)
## histogram of the fully observed hazard data (gray) and censored observations (light blue)
hist( SimDat$time[ which(SimDat$status==1)], main="", xlab="t", breaks="Freedman-Diaconis", xlim=c(0,10) )
hist( SimDat$time[ which(SimDat$status==0)], main="", xlab="t", breaks="Freedman-Diaconis", xlim=c(0,10),add=TRUE, col=rgb(173,216,230,max=255,alpha=100) )

```

### Hazard regression

```{r regression}
## Fit a cubic B-spline regression model using the true knots
   yd0= read.csv("C:/local/CORU/Survival Hazard and KM/Philip Rosenberg papers and code/matlab v1 PR/Dataset1.csv", header = F, col.names = c("time","status") )
   SimDat = yd0
   timeout = seq( 0, 10, length.out = 101 )
   Result = hspcore(yd=SimDat, ORDER=4, knots=c(0,1,3,6,10), time=timeout, Bootstrap = 200 )   
```

You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this.

You can also embed plots, for example:

```{r pressure, echo = FALSE}
plot(pressure)
```

In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN.
